---
title: "Drug repurposing for breast cancer treatment - team04"
author: "Simon Westermann, Luis Herfurth, Aaron Eidenmüller, Sharujan Suthakaran"
date: "29.April 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Information about the data

I just loaded the rda files in here. If you moved the rda files and this rmd in the same working directory, than the load command will work for you aswell. 

```{r}
load("prism_datasets.rda")
load("celllines_datasets.rda")

```

### Prism Datasets  

**Prism:** effect of the treatment (columns) on cell growth of the cell lines (rows); includes drug, dosage and assay used

**Prism.treat:** for each treatment (rows) further information on the treatment and drug

**Prism.cl:** contains information about the different celllines    
*if we search after "breast" in the column lineage we get our 22 celllines*

### Cellline Datasets  

**Prism.exp:** contains levels of gene expression. Celllines (rows) and genes (columns)

**Prism.cnv:** contains copy number levels of genes. Normal is CN = 2. Gene names (rows) and celllines (columns)

**Prism.snv:** marks mutation in the different celllines als functional or nonfunctional to the cancer.

**Prism.achilles:** has information on how important a gene is for cell survival. Was generated using knockdown celllines. Gene names (rows) and celllines (colums)
   
   
## Clean up of the data
