```{r}
load("C:/Users/aaron/Desktop/Studium/4. Semester/Bioinfo/prism_datasets.rda")
load("C:/Users/aaron/Desktop/Studium/4. Semester/Bioinfo/cellline_datasets.rda")

```

```{r}
sum(is.na(prism.achilles))
dim(prism.achilles)

```
```{r}
sum(apply(prism.achilles, 1, function(x){any(is.na(x))})) #3
sum(apply(prism.achilles, 2, function(x){any(is.na(x))})) #1936

#Only three cell lines have NA values. But which?

na.celllines <- which(apply(prism.achilles, 1, function(row) any(is.na(row))))
#ACH-000836 ACH-000901 ACH-000939 
   #    277        309        323 
prism.cl[na.celllines,]
#SUIT2_PANCREAS pancreas cancer
#TCCSUP_URINARY_TRACT urinary tract cancer
#NCIH1703_LUNG lung cancer

#Which percentage of values are NA values?
na.amount <- apply(prism.achilles[na.celllines,],1, function(x) {
  sum(is.na(x))
})
(na.amount)/(ncol(prism.achilles))

#ACH-000836 ACH-000901 ACH-000939 
#0.07163751 0.03896462 0.07163751

```
The percentage is about 3%-7% of all genes for these cell lines.
To investigate how to imputate we will have to look at the median and mean

```{r}
apply(prism.achilles[na.celllines,],1 ,function(x){
mean(x,na.rm=TRUE)  
} )
#ACH-000836 ACH-000901 ACH-000939 
#-0.1623019 -0.1582249 -0.1437818
apply(prism.achilles[na.celllines,],1 ,function(x){
median(x,na.rm=TRUE)  
} )
#ACH-000836  ACH-000901  ACH-000939 
#-0.06607414 -0.06220032 -0.05627875 

```
The median and mean differ. We try to imputate the median as the NA values
```{r}

prism.achilles.rm <-apply(prism.achilles, 2, function(x) {
  x[which(is.na(x))] <- median(x, na.rm = TRUE)
  return(x)
})


```
Now we will calculate the mean, median, standard deviation, minimum and maximum of each gene
```{r}
prism.achilles.mean <- apply(prism.achilles.rm, 2, mean)

prism.achilles.median <- apply(prism.achilles.rm, 2, median)

prism.achilles.sd <- apply(prism.achilles.rm, 2, sd)

prism.achilles.min <- apply(prism.achilles.rm, 2, min)

prism.achilles.max <- apply(prism.achilles.rm, 2, max)
```
Now we try to put all of this information in a data frame

```{r}
prism.achilles.overview <- cbind(prism.achilles.mean, prism.achilles.median, prism.achilles.sd, prism.achilles.min, prism.achilles.max)
colnames(prism.achilles.overview) <- c("mean", "median", "standard deviation", "minimum", "maximum")
```
The data frame "prism.achilles.overview" now contains all the above mentioned characteristics. Now we will look at the histogramms.

```{r}
hist(prism.achilles.overview[,1], main = "mean of all the prism.achilles genes", xlab = "mean")

hist(prism.achilles.overview[,2], main = "median of all the prism.achilles genes", xlab = "median")

hist(prism.achilles.overview[,3], main = "standard deviation of all the prism.achilles genes", xlab = "standard deviation")

hist(prism.achilles.overview[,4], main = "minimum of all the prism.achilles genes", xlab = "minimum")

hist(prism.achilles.overview[,5], main = "maximum of all the prism.achilles genes", xlab = "maximum")
```


```{r}
breast.cl <- which(prism.cl[,"lineage"] == "breast")
prism.achilles.breast <- prism.achilles.rm[rownames(prism.achilles.rm) %in% prism.cl[breast.cl, 1], ]
library(pheatmap)
pheatmap(prism.achilles.breast, show_colnames = FALSE)
```















