# 1. damaging Genes

***integrate prism.snv for easier gene/cell line filter***

```{r}

PRISM.snv.breast <- prism.snv[which(as.character(prism.snv$DepMap_ID) %in% breast_prism.cl$DepMap_ID),]

prism.snv.breast <- PRISM.snv.breast[PRISM.snv.breast$Variant_Classification != "Silent",]

```

***I will try to integrate via subtypes in order to get possible treatments for best respone***

```{r}
###manuell clustering

breast_carcinoma <- c(breast_prism.cl[which(breast_prism.cl$lineage_subtype == "breast_carcinoma"),])$DepMap_ID

breast_ductal_carcinoma <- c(breast_prism.cl[which(breast_prism.cl$lineage_subtype == "breast_ductal_carcinoma"),])$DepMap_ID

#breast_adenocarcinoma <- c(breast_prism.cl[which(breast_prism.cl$lineage_subtype == "breast_adenocarcinoma"),])$DepMap_ID

```

```{r}

prism.achilles.breast_carcinoma <- prism.achilles.breast[which(rownames(prism.achilles.breast)%in%breast_carcinoma),]

prism.achilles.breast_ductal_carcinoma <- prism.achilles.breast[which(rownames(prism.achilles.breast)%in%breast_ductal_carcinoma),]

```

*focus on damaging variants*

```{r}

prism.snv.breast_damage <- prism.snv.breast[prism.snv.breast$Variant_annotation == "damaging",]

##for breast carcinoma

prism.snv.breast_carcinoma <- prism.snv.breast_damage[prism.snv.breast_damage$DepMap_ID %in% breast_carcinoma,]

##for breast ductal carcinoma

prism.snv.breast_ductal_carcinoma <- prism.snv.breast_damage[prism.snv.breast_damage$DepMap_ID %in% breast_ductal_carcinoma,]

```

**prism.achilles does not have every breast cancer cell line, damage dataframes**

```{r}

missingrows.achilles <- matrix(0, nrow = 5, ncol = ncol(prism.achilles.breast))

colnames(missingrows.achilles) <- colnames(prism.achilles.breast)

rownames(missingrows.achilles) <- setdiff(rownames(breast_prism.exp), rownames(prism.achilles.breast))[1:5]

breast_prism.achilles <- rbind(prism.achilles.breast, missingrows.achilles)

dim(breast_prism.achilles)


###damage gene data frame for .exp

damage.exp <- breast_prism.exp[,colnames(breast_prism.exp) %in% prism.snv.breast_damage$Hugo_Symbol]

exp.sorted_col <- colnames(damage.exp)[order(colnames(damage.exp))]

breast_damage.exp <- damage.exp[, exp.sorted_col]


###damage gene data frame for .cnv

damage.cnv <- breast_prism.cnv[,colnames(breast_prism.achilles) %in% prism.snv.breast_damage$Hugo_Symbol]

cnv.sorted_col <- colnames(damage.cnv)[order(colnames(damage.cnv))]

breast_damage.cnv <- damage.cnv[, cnv.sorted_col]

###damage gene data frame for .achilles

breast_damage.achilles <- as.data.frame(breast_prism.achilles[,colnames(breast_prism.achilles) %in% prism.snv.breast_damage$Hugo_Symbol])

damage.achilles_mean <- apply(breast_damage.achilles, 2, mean)

```

**damage genes in .exp and .achilles for lineage subtypes**

```{r}

carcinoma_damage.achilles <- as.data.frame(breast_prism.achilles[rownames(prism.achilles.breast_carcinoma),colnames(breast_prism.achilles) %in% prism.snv.breast_damage$Hugo_Symbol])

carcinoma.damage.achilles_mean <- apply(carcinoma_damage.achilles, 2, mean)

#################

important_carcinoma.genes <- colnames(carcinoma_damage.achilles[,which(carcinoma.damage.achilles_mean <= -0.5)])

important_carcinoma.gene_exp <- prism.exp[, colnames(prism.exp) %in% important_carcinoma.genes]

important_carcinoma.gene_list <- colnames(important_carcinoma.gene_exp)

```

```{r}

ductal_carcinoma_damage.achilles <- as.data.frame(breast_prism.achilles[rownames(prism.achilles.breast_ductal_carcinoma),colnames(breast_prism.achilles) %in% prism.snv.breast_damage$Hugo_Symbol])

ductal_carcinoma.damage.achilles_mean <- apply(ductal_carcinoma_damage.achilles, 2, mean)

########################

important_ductal.genes <- colnames(ductal_carcinoma_damage.achilles[,which(ductal_carcinoma.damage.achilles_mean <= -0.5)])

important_ductal.gene_exp <- prism.exp[, colnames(prism.exp) %in% important_ductal.genes]

important_ductal.gene_list <- colnames(important_ductal.gene_exp)

```

#2 Overview for loop

```{r}

for (gene_name in prism.snv.breast_damage$Hugo_Symbol) {
   
    matchingRows <- which(prism.treat_subset$target == gene_name)
    if (length(matchingRows) == 0) {
      cat("No matching treatment found for Gene:", gene_name, "\n")
    } else {
      cat("Matching treatment(s) for Gene:", gene_name, "\n")
      for (row in matchingRows) {
        rowName <- row.names(prism.treat_subset)[row]
        rowValue <- as.character(prism.treat_subset$name[row])
        cat("Row Name:", rowName, "\tTreatment:", rowValue, "\n")
      }
    }
    
    
    if (gene_name %in% colnames(breast_damage.exp) &&
        gene_name %in% colnames(breast_damage.achilles) &&
        gene_name %in% colnames(breast_damage.cnv)) {
      
      
      exp_value <- breast_damage.exp[rownames(breast_damage.exp), gene_name]
      cat("Expression Level:", exp_value, "\n")
      
      
      achilles_value <- breast_damage.achilles[rownames(breast_damage.achilles), gene_name]
      cat("Achilles Value:", achilles_value, "\n")
      
    
      cnv_value <- breast_damage.cnv[rownames(breast_damage.cnv), gene_name]
      cat("Copy Number:", cnv_value, "\n")
      
      cat("\n")
    } else {
      cat("Gene:", gene_name, "is not present in all datasets.\n\n")
      
    if (gene_name %in% colnames(breast_damage.exp)) {
      exp_value <- breast_damage.exp[, gene_name]
      cat("Expression Level (breast_damage.exp):\n")
      for (i in 1:length(rownames(breast_damage.exp))) {
      cat(rownames(breast_damage.exp)[i], ":", exp_value[i], "\n")
}
    }
    
    if (gene_name %in% colnames(breast_damage.achilles)) {
      achilles_value <- breast_damage.achilles[, gene_name]
      cat("Achilles Value (breast_damage.achilles):\n")
      for (j in 1:length(rownames(breast_damage.achilles))) {
      cat(rownames(breast_damage.achilles)[j], ":", achilles_value[j], "\n")
}
    }
    
    if (gene_name %in% colnames(breast_damage.cnv)) {
      cnv_value <- breast_damage.cnv[, gene_name]
      cat("Copy Number (breast_damage.cnv):\n")
      for (z in 1:length(rownames(breast_damage.cnv))) {
      cat(rownames(breast_damage.cnv)[z], ":", cnv_value[z], "\n")
}
    }
    
    cat("\n")
  }
}

```

***Search engine for treatments, needs further improvement***

```{r}

treatmentSearchEngine <- function(y) {
  matchingRows <- which(prism.treat$name == y)
  if (length(matchingRows) == 0) {
    cat("No matching target found.\n")
  } else {
    cat("Matching target(s):\n")
    for (row in matchingRows) {
      rowName <- row.names(prism.treat)[row]
      rowValue <- as.character(prism.treat$target[row])
      cat("Row Name:", rowName, "\tTarget:", rowValue, "\n")
    }
  }
}


treatmentSearchEngine("noretynodrel")

```