---
title: "Drug repurposing for breast cancer treatment - team04"
author: "Simon Westermann, Luis Herfurth, Aaron Eidenmüller, Sharujan Suthakaran"
date: "29.April 2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Information about the data

I just loaded the rda files in here. If you moved the rda files and this rmd in the same working directory, than the load command will work for you aswell.

```{r}
load("~/Downloads/prism_datasets.rda")
load("~/Downloads/cellline_datasets.rda")


library(knitr)
library(rmdformats)
library(pheatmap)

```

### Prism Datasets

**Prism:** effect of the treatment (columns) on cell growth of the cell lines (rows); includes drug, dosage and assay used

**Prism.treat:** for each treatment (rows) further information on the treatment and drug

**Prism.cl:** contains information about the different celllines\
*if we search after "breast" in the column lineage we get our 22 celllines*

### Cellline Datasets

**Prism.exp:** contains levels of gene expression. Celllines (rows) and genes (columns)

**Prism.cnv:** contains copy number levels of genes. Normal is CN = 2. Gene names (rows) and celllines (columns).pr

**Prism.snv:** marks mutation in the different celllines als functional or nonfunctional to the cancer.

**Prism.achilles:** has information on how important a gene is for cell survival. Was generated using knockdown celllines. Gene names (rows) and celllines (colums)

## Clean up of the data

```{r}

#View(prism.exp)

class(prism.exp)

#plot(prism.exp) #too many numbers for R to plot

#dimension and column number available in environment

getOption("max.print") 

dim(prism.exp)


```

```{r}
####general info
min(prism.exp, na.rm = T) # 0
max(prism.exp, na.rm = T) # 17.25397

###cleaning NAs
sum(is.na(prism.exp)) #non or NA=0, no expression

### checking for zeros instead of NA
sum(which(prism.exp == 0)) #8,2244 x 10^12 zeros = not detected

deviation <- apply(prism.exp, 2, function(x) sum(x == 0)/length(x))

mean(apply(prism.exp, 2, function(x) sum(x == 0)/length(x))) #0.1531 => 15,3 % of each column is zero

zeros.exp<- mean(apply(prism.exp, 2, function(x) sum(x == 0)/length(x)))

deviation.exp <- deviation - zeros.exp

hist(deviation.exp, breaks = 20)

colzero.exp <- apply(prism.exp, 2, function(x) all(x == 0))

which(colzero.exp) # 6 Genes not expressed in any cell

#--> delete these columns? Yes

PRISM.exp <- prism.exp[, !colzero.exp]

#View(PRISM.exp)

```

```{r}
###breast cancer <> gene expression

#View(prism.cl)

breast_prism.exp <- prism.exp[which(rownames(prism.exp) %in% lineage.breast$DepMap_ID),]

lineage.breast <- prism.cl[which(prism.cl$lineage == "breast"),]

#View(breast_prism.exp)

##pheatmap(breast_prism.exp, show_rownames = F, show_colnames = F) # many Genes not detected (middle)

sum(which(breast_prism.exp == 0))

mean(apply(breast_prism.exp, 2, function(x) sum(x == 0)/length(x))) #0.1549 => 15,5 % of each column is zero

breastzero.col <- apply(breast_prism.exp, 2, function(x) all(x == 0))

sum(breastzero.col) # 449 Genes are not expressed in breast cancer cells

#--> delete these columns?

breast_PRISM.exp <- breast_prism.exp[, !breastzero.col]

#View(breast_PRISM.exp)


#pheatmap(breast_PRISM.exp, show_rownames = T, show_colnames = F)

```

```{r}
#### next: correlation between cell lines (expressing same genes?), clustering by genome similarities

##correlation between genes
breast_PRISMexp.cor <- cor(breast_PRISM.exp, method = "pearson", use = "pairwise.complete.obs")

dim(breast_PRISMexp.cor) # 18728 x 18728

pheatmap(breast_PRISMexp.cor, show_rownames = F , show_colnames = F) #too big

##kmeans

wss <- sapply(1:10, function(x) kmeans(breast_PRISM.exp, x)$tot.withinss)

plot(1:10, wss, type = "b", xlab = "N of clusters", ylab = "Total within sum of square") #very strange hahah

```

```{r}
##pca without thinking
pca <- prcomp(breast_PRISM.exp, center = T, scale. = T)

summary(pca)

#View(pca)

plot(pca$x[,1],pca$x[,2], col= breast_PRISM.exp, pch=19 ,xlab = "PC1", ylab = "PC2") 

#View(prism.achilles) 


class(breast_prism.exp)
```
```{r}
# Perform PCA
pca <- prcomp(breast_PRISM.exp, center = TRUE, scale. = TRUE)

# Summary of PCA results
summary(pca)

# Plot PCA
plot(pca$x[, 1], pca$x[, 2], col = "blue", pch = 19, xlab = "PC1", ylab = "PC2")

# Add cell line labels
text(pca$x[, 1], pca$x[, 2], pos = 4, cex = 0.8)



```

```{r}
# Assign cell line lineage from prism.cl to breast_PRISM.exp
lineages <- prism.cl$lineage[rownames(breast_PRISM.exp)]

# Create a vector of colors for the lineages
lineage_colors <- rainbow(length(unique(lineages)))

# Plot PCA with lineage colors
plot(pca$x[, 1], pca$x[, 2], col = lineage_colors[match(lineages, unique(lineages))], pch = 19, xlab = "PC1", ylab = "PC2")

# Add legend for lineages
legend("bottomright", legend = unique(lineages), col = lineage_colors, pch = 19, cex = 0.8)

```



```{r}
#k-means prism.exp mit Varianz

breast_sd.exp <- apply(breast_prism.exp, 2, sd)

min(breast_sd.exp)

max(breast_sd.exp)

hist(breast_sd.exp)

test.exp <- breast_prism.exp[, which(breast_sd.exp > 2)]

#View(breast_big1.exp)

#View(test.exp)

wss.exp <- sapply(1:10, function(x) kmeans(test.exp, x)$tot.withinss)

plot(1:10, wss.exp, type = "b", xlab = "N of clusters", ylab = "Total within sum of square")

```

```{r}
#silhouette

library(cluster)

silhouette(test.integer) #does not work yet

sum(is.na(breast_prism.exp))

class(breast_prism.exp)

test.integer <- apply(breast_prism.exp, 2, as.integer)

rm(test.integer)


D = dist(t(breast_prism.exp))

km = kmeans(x=t(breast_prism.exp), centers = 2, nstart = 10)

s = silhouette(km$cluster,D)

plot(s)

```